---
title: "P8131 Assignment 2"
author: "Yihan Feng (yf2555)"
date: "2/9/2021"
output: html_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library("ResourceSelection")
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F)
```

<br>

# Problem 1

```{r, include = FALSE}
dose = 0:4

dying_alive = tibble(
  dying_n = c(2,8,15,23,27),
  alive_n = 30 - dying_n
) %>%
  as.matrix()
```

<br>

### a. Fill out the table and give comments. 

```{r}
# fit models
fit_logit = glm(dying_alive ~ dose, family = binomial(link = 'logit'))
fit_probit = glm(dying_alive ~ dose, family = binomial(link = 'probit'))
fit_clog = glm(dying_alive ~ dose, family = binomial(link = 'cloglog'))

# CI for beta
ci_logit = fit_logit %>%
  broom::tidy() %>%
 mutate(lower = estimate - 1.96 * std.error,
        upper = estimate + 1.96 * std.error) %>%
 select(term, estimate, lower, upper) %>%
 filter(term == "dose")

ci_probit = fit_probit %>%
 broom::tidy() %>%
 mutate(lower = estimate - 1.96 * std.error,
        upper = estimate + 1.96 * std.error) %>%
 select(term, estimate, lower, upper) %>%
 filter(term == "dose")
 
ci_clog = fit_clog %>%
 broom::tidy() %>%
 mutate(lower = estimate - 1.96 * std.error,
        upper = estimate + 1.96 * std.error) %>%
 select(term, estimate, lower, upper) %>%
 filter(term == "dose")
 
# p(dying|x=0.01)
new_dose <- data.frame(dose = 0.01)

predict_logit = predict(fit_logit, new_dose, type = "response")
predict_probit = predict(fit_probit, new_dose, type = "response")
predict_clog = predict(fit_clog, new_dose, type = "response")

# set up the table
table_a = tibble(
 model = c("Logit", "Probit", "C-log-log"),
 beta_estimate = c(fit_logit$coefficients[2], fit_probit$coefficients[2], fit_clog$coefficients[2]),
 beta_cI_lower = c(ci_logit$lower, ci_probit$lower, ci_clog$lower),
 beta_CI_upper = c(ci_logit$upper, ci_probit$upper, ci_clog$upper),
 Deviance = c(fit_logit$deviance, fit_probit$deviance, fit_clog$deviance),
 p_dying_x_0.01 = c(predict_logit, predict_probit, predict_clog)
) %>%
  knitr::kable(digits = 3)

table_a
```


According to the table, I noticed that: 

* All estimates of $\beta$ are positive, which means that the risk of dying increases with the increase of doses. The logit model has the highest value, and the probit model has the lowest value. 

* All confidence intervals of $\beta$ do not include zero, which corresponds to the conclusion that the risk of dying increases with the increase of doses. 

* The c-log-log model has the highest deviance, and the probit model has the lowest deviance. It indicates that the logit and probit may better fit the data set. 

* When the dose is decreased to 0.01, the three models provide similar probabilities of dying. 


<br>
<br>

### b. Suppose that the dose level is in natural logarithm scale, estimate LD50 with 90% confidence interval

```{r}
#logit model

logit_beta0 = fit_logit$coefficients[1]
logit_beta1 = fit_logit$coefficients[2]
logit_betacov = vcov(fit_logit)
logit_x0fit = -logit_beta0 / logit_beta1
logit_estimate = exp(logit_x0fit)
logit_varx0 = logit_betacov[1,1] / (logit_beta1^2) + logit_betacov[2,2] * (logit_beta0^2) / (logit_beta1^4) - 2 * logit_betacov[1,2] * logit_beta0/(logit_beta1^3)
logit_CI = exp(logit_x0fit + c(qnorm(0.05), -qnorm(0.05)) * sqrt(logit_varx0))
 

#probit model

probit_beta0 = fit_probit$coefficients[1]
probit_beta1 = fit_probit$coefficients[2]
probit_betacov = vcov(fit_probit)
probit_x0fit = -probit_beta0 / probit_beta1
probit_estimate = exp(probit_x0fit) 
probit_varx0 = probit_betacov[1,1] / (probit_beta1^2) + probit_betacov[2,2] * (probit_beta0^2) / (probit_beta1^4) - 2 * probit_betacov[1,2] * probit_beta0 / (probit_beta1^3)
probit_CI = exp(probit_x0fit + c(qnorm(0.05), -qnorm(0.05)) * sqrt(probit_varx0))


#cloglog model

clog_beta0 = fit_clog$coefficients[1]
clog_beta1 = fit_clog$coefficients[2]
clog_betacov = vcov(fit_clog)
clog_x0fit = -clog_beta0 / clog_beta1
clog_estimate = exp(clog_x0fit) 
clog_varx0 = clog_betacov[1,1] / (clog_beta1^2) + clog_betacov[2,2] * (clog_beta0^2) / (clog_beta1^4) - 2 * clog_betacov[1,2] * clog_beta0 / (clog_beta1^3)
clog_CI = exp(clog_x0fit + c(qnorm(0.05), -qnorm(0.05)) * sqrt(clog_varx0))


#set up table
table_b = tibble(
 model = c("Logit", "Probit", "C-log-log"),
 beta_estimate = c(logit_estimate, probit_estimate, clog_estimate),
 beta_cI_lower = c(logit_CI[1], probit_CI[1], clog_CI[1]),
 beta_CI_upper = c(logit_CI[2], probit_CI[2], clog_CI[2])
) %>%
  knitr::kable(digits = 3)

table_b
```


<br>
<br>
<br>
<br>

## Problem 2

<br>

### a. How does the model fit the data? 

```{r}
mph_df = data.frame(amount = seq(from = 10, to = 90, by = 5), 
                      offers = c(4, 6, 10, 12, 39, 36, 22, 14, 10, 12, 8, 9, 3, 1, 5, 2, 1), 
                      enrolls = c(0, 2, 4, 2, 12, 14, 10, 7, 5, 5, 3, 5, 2, 0, 4, 2, 1))

resp = cbind(mph_df$enrolls, mph_df$offers - mph_df$enrolls)

mph_fit = glm(resp ~ mph_df$amount, family = binomial(link = 'logit'))
mph_hl = hoslem.test(mph_fit$y, fitted(mph_fit), g = 10)

mph_hl
```

From the Hosmer-Lemeshow test, the p value is 0.991 ( >0.05). Therefore, it fails to reject, and the model fit the data well. 

<br>
<br>

### b. How do you interpret the relationship between the scholarship amount and enrollment rate? What is 95% CI? 

```{r}
beta_mph = mph_fit$coefficients[2]

ci_mph = mph_fit %>%
  broom::tidy() %>%
  mutate(lower = beta_mph - 1.96 * std.error,
        upper = beta_mph + 1.96 * std.error) %>%
 select(term, estimate, lower, upper) %>%
 filter(term == "mph_df$amount") %>%
  knitr::kable(digits = 3)

ci_mph
```

<br>
<br>

### c. How much scholarship should we provide to get 40% yield rate? What is the 95% CI? 

```{r}
mph_beta0 = mph_fit$coefficients[1]
mph_beta1 = mph_fit$coefficients[2]

mph_betacov = vcov(mph_fit) 

mph_x0fit = (log(0.4/0.6) - mph_beta0) / mph_beta1

mph_varx0 = mph_betacov[1,1]/(mph_beta1^2) + mph_betacov[2,2]*((mph_beta0 - log(2/3))^2)/(mph_beta1^4) -
              2*mph_betacov[1,2]*(mph_beta0 - log(2/3))/(mph_beta1^3)


ci_mph1 = mph_fit %>%
  broom::tidy() %>%
  mutate(lower = mph_x0fit + qnorm(0.025) * sqrt(mph_varx0),
        upper = mph_x0fit - qnorm(0.025) * sqrt(mph_varx0)) %>%
 select(term, estimate, lower, upper) %>%
 filter(term == "mph_df$amount") %>%
  knitr::kable(digits = 3)

ci_mph1
```

We need `r round(mph_x0fit, 3)` thousands of dollars scholarship to get 40% yield rate. 

